---
title: "Antimicrobial Resistance Draft"
author: "Arvon Clemons II, Ashika Mani, Mark Vater, Valerio Tonelli"
date: "9/27/2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE, eval=FALSE}
library(tidyverse)
# read metadata.csv with only "Sample_ID" and "*_sr" features
metadata <- readr::read_csv("./metadata.csv") %>% 
  select(Sample_ID,(matches("_sr$")))

# read antibiotic resistance associated unitigs
unitigs_azm <- read_table2("azm_sr_gwas_filtered_unitigs.Rtab")
unitigs_cfx <- read_table2("cfx_sr_gwas_filtered_unitigs.Rtab")
unitigs_cip <- read_table2("cip_sr_gwas_filtered_unitigs.Rtab")

#Function to transpose unitigs table, join metadata, then create .csv file
unitigs <- function(unitig){
  
  # deparse + substitute combo to extract object name
  unitig_name <- deparse(substitute(unitig))
  
  '%>%' <- tidyr::'%>%'
  
  unitig %>%
    # transpose
    tidyr::pivot_longer(-pattern_id, "Sample_ID", "value") %>% 
    tidyr::pivot_wider(Sample_ID, pattern_id) %>%
    # join relevant antibiotic metadata
    dplyr::inner_join(dplyr::select(metadata, "Sample_ID", dplyr::matches(stringr::str_extract(unitig_name, "[[:alpha:]]+$"))), by = "Sample_ID") %>% 
    dplyr::select(dplyr::last_col(), dplyr::everything(), -"Sample_ID") %>% 
    # rename unitigs to format "geneX" where X is a positive integer
    data.table::setnames(., old = names(.[2:length(.)]), new = paste0('gene', seq_along(.[2:length(.)])), skip_absent = T) %>%
    # relabel into "neg" and "pos"
    dplyr::mutate(across(where(is.numeric), ~if_else(. == 0, "neg", "pos"))) %>% 
    tidyr::drop_na() %>% 
    # create .csv file
    readr::write_csv(path=paste0('./', stringr::str_extract(unitig_name,"[[:alpha:]]+[_][[:alpha:]]+"), '.csv'))
}

# Use 'unitigs' function to create .csv files
unitigs(unitigs_azm)
unitigs(unitigs_cfx)
unitigs(unitigs_cip)
```

```{r, include=FALSE}
#libraries
library(caret)
library(dplyr)
library(ggpubr)

# Create list of models
RF <- lapply(list.files(pattern = "RF_model"), readRDS) %>%
  setNames(na.omit((stringr::str_extract(list.files(), pattern = "[:alpha:]+RF_model$"))))

EN <- lapply(list.files(pattern = "EN_model"), readRDS) %>%
  setNames(na.omit((stringr::str_extract(list.files(), pattern = "[:alpha:]+EN_model$"))))

SVM <- lapply(list.files(pattern = "SVM_model"), readRDS) %>%
  setNames(na.omit((stringr::str_extract(list.files(), pattern = "[:alpha:]+SVM_model$"))))

XGB <- lapply(list.files(pattern = "XGB_model"), readRDS) %>%
  setNames(na.omit((stringr::str_extract(list.files(), pattern = "[:alpha:]+XGB_model$"))))

# createDS()
createDS <- function(csvname, percent=0.7, seed=05101929){
  stopifnot(is.character(csvname))
  
  '%>%' <- tidyr::'%>%'
  
#Import and create Training/Testing datasets
rfData <- readr::read_csv(csvname, col_types = readr::cols(.default = readr::col_factor(NULL))) %>% 
  dplyr::select_if(sapply(., nlevels) > 1)
#Split
set.seed(seed)
index <- dplyr::select(rfData,dplyr::matches(stringr::str_extract(unlist(stringr::str_split(csvname, "_"))[2],"^[[:alpha:]]+"))) %>% 
  unlist() %>% 
  caret::createDataPartition(times = 1, p = percent, list = FALSE)

dfTrain <- rfData %>% dplyr::slice(index) #training dataset
dfTest <- rfData %>% dplyr::slice(-index) #testing dataset

#return datasets
dataSets <- list(training = dfTrain, testing = dfTest)
invisible(dataSets)
}
azmDS <- createDS("./unitigs_azm.csv")
cfxDS <- createDS("./unitigs_cfx.csv")
cipDS <- createDS("./unitigs_cip.csv")
```
# Introduction

Antibiotic resistance is the property of bacteria to resist the effects of antibiotic compounds, which occurs when resistant bacteria survive exposure and continue to multiply.1 This occurrence is not only a large financial burden but poses a large hazard to human health as well. According to the Centers for Disease Control and Prevention (CDC), the economic impact of antibiotic resistance may account for as much as $20 billion in direct healthcare costs and \$35 billion in associated consequences annually in the US.2 Furthermore, there are at least 23,000 Antibiotic resistant-related deaths occurring each year as well as at least 2,000,000 illnessess.2 Therefore, this issue is of vital importance from both a public health and financial standpoint. 

In this analysis, we will use a dataset that has been assembled from a variety of publications on bacterial resistance and is freely available on Kaggle. We will assess the genetic variations associated with resistance to three different antibiotics that are widely used for the treatment of gonorrhea (azithromycin, ciprofloxacin, and cefixime). These segments of DNA are known as unitigs and are shared by the Neisseria gonorrhoeae strains that are included in this dataset. The dataset has about 3000 data entries for each antibiotic tested, with one having approximately 300 unitigs, a second one having approximately 500 unitigs, and the last one having over 3000 unitigs. We will be looking at using various statistical approaches to create machine learning algorithms to best predict the genetic background that determines resistance to these antibiotics. The models that we anticipate using will include elastic net logistic regression, XGBoost, random forest, and support vector machine. Prediction models outcomes will then be tested for their performance to draw conclusions from the comparative analyses.

# Data Description

A dataset of multiple unitig sequences associated with antibiotic resisistance in *Neisseria gonorrhea*, outcomes are classified upon whether resistance has been identified into two classes ("Positive", "Negative).
```{r, echo=FALSE}
knitr::kable(data.frame(Observations = c(nrow(azmDS$training) + nrow(azmDS$testing), nrow(cfxDS$training) + nrow(cfxDS$testing), nrow(cipDS$training) + nrow(cipDS$testing)),
           Features = c(ncol(azmDS$training) - 2, ncol(cfxDS$training) - 2, ncol(cipDS$training) - 2),
           row.names = c("AZM", "CFX", "CIP")
           ), align = c("c", "c"), caption = "Data Description")
```


# Data Analysis

Four supervised classification methods were used: (1) Elastic Net(EN), (2) Support Vector Machine(SVM), (3) Extreme Gradient Boosting(XGB), (4) Random Forest(RF). All methods were implemented through the `caret` `r packageVersion("caret")` package. We used ROC as the metric for selecting the optimal threshold in each model.

## Elastic Net

### Azithromycin

#### Training

```{r, echo = FALSE}
confusionMatrix.train(EN$azmEN_model)
knitr::kable(EN$azmEN_model$bestTune, caption = "AZM Elastic Net tuned parameters")
```

#### Final Model Assessment

```{r, echo=FALSE}
predictedAR <- predict(EN$azmEN_model, azmDS$testing) #full predictions
knitr::kable(confusionMatrix(predictedAR, azmDS$testing$azm_sr)$table, caption = "Azithromycin Elastic Net Confusion Matrix") #confusion matrix
knitr::kable(confusionMatrix(predictedAR, azmDS$testing$azm_sr)$byClass[c(1:4, 11)], caption = "Azithromycin Elastic Net Perfomance Statistics") # Sensitivity, Specificity, Pos Pred Value, Neg Pred value, Balanced Accuracy
azmEN_tune <- plot.train(EN$azmEN_model) #graphs Tuning Accuracy
azmEN_Imp <- varImp(EN$azmEN_model) %>% dotPlot() #graphs Importance
```

### Cefoxitin

#### Training

```{r, echo = FALSE}
confusionMatrix.train(EN$cfxEN_model)
knitr::kable(EN$cfxEN_model$bestTune, caption = "CFX Elastic Net tuned parameters")
```

#### Final Model Assessment

```{r, echo=FALSE}
predictedAR <- predict(EN$cfxEN_model, cfxDS$testing) #full predictions
knitr::kable(confusionMatrix(predictedAR, cfxDS$testing$cfx_sr)$table, caption = "Cefoxitin Elastic Net Confusion Matrix") #confusion matrix
knitr::kable(confusionMatrix(predictedAR, cfxDS$testing$cfx_sr)$byClass[c(1:4, 11)], caption = "Cefoxitin Elastic Net Performance Statistics") # Sensitivity, Specificity, Pos Pred Value, Neg Pred value, Balanced Accuracy
cfxEN_tune <- plot.train(EN$cfxEN_model) #graphs Tuning Accuracy
cfxEN_Imp <- varImp(EN$cfxEN_model) %>% dotPlot() #graphs Importance
```


### Ciprofloxacin

#### Training

```{r, echo = FALSE}
confusionMatrix.train(EN$cipEN_model)
knitr::kable(EN$cipEN_model$bestTune, caption = "CIP Elastic Net tuned parameters")
```

#### Final Model Assessment

```{r, echo=FALSE}
predictedAR <- predict(EN$cipEN_model, cipDS$testing) #full predictions
knitr::kable(confusionMatrix(predictedAR, cipDS$testing$cip_sr)$table, caption = "Ciprofloxacin Elastic Net Confusion Matrix") #confusion matrix
knitr::kable(confusionMatrix(predictedAR, cipDS$testing$cip_sr)$byClass[c(1:4, 11)], caption = "Ciprofloxacin Elastic Net Performance Statistics") # Sensitivity, Specificity, Pos Pred Value, Neg Pred value, Balanced Accuracy
cipEN_tune <- plot.train(EN$cipEN_model) #graphs Tuning Accuracy
cipEN_Imp <- varImp(EN$cipEN_model) %>% dotPlot() #graphs Importance
```

### Hyperparameter Tuning Comparison

```{r, echo = FALSE, fig.cap = "ROC vs Regularization Tuning: AZM, CFX, and CIP from top to bottom"}
gridExtra::grid.arrange(azmEN_tune, cfxEN_tune, cipEN_tune)
```

## Support Vector Machines

### Azithromycin

#### Training

```{r, echo = FALSE}
confusionMatrix.train(SVM$azmSVM_model)
knitr::kable(SVM$azmSVM_model$bestTune, caption = "AZM Support Vector Machines tuned parameters")
```

#### Final Model AssessmSVMt

```{r, echo=FALSE, error = TRUE}
predictedAR <- predict(SVM$azmSVM_model, azmDS$testing) #full predictions
knitr::kable(confusionMatrix(predictedAR, azmDS$testing$azm_sr)$table, caption = "Azithromycin Support Vector Machines Confusion Matrix") #confusion matrix
knitr::kable(confusionMatrix(predictedAR, azmDS$testing$azm_sr)$byClass[c(1:4, 11)], caption = "Azithromycin Support Vector Machines Performance Statistics") # Sensitivity, Specificity, Pos Pred Value, Neg Pred value, Balanced Accuracy
azmSVM_tune <- plot.train(SVM$azmSVM_model) #graphs Tuning Accuracy
#azmSVM_Imp <- varImp(SVM$azmSVM_model) %>% dotPlot() #graphs Importance
```

### Cefoxitin

#### Training

```{r, echo = FALSE}
confusionMatrix.train(SVM$cfxSVM_model)
knitr::kable(SVM$cfxSVM_model$bestTune, caption = "CFX Support Vector Machines tuned parameters")
```

#### Final Model AssessmSVMt

```{r, echo=FALSE}
predictedAR <- predict(SVM$cfxSVM_model, cfxDS$testing) #full predictions
knitr::kable(confusionMatrix(predictedAR, cfxDS$testing$cfx_sr)$table, caption = "Cefoxitin Support Vector Machines Confusion Matrix") #confusion matrix
knitr::kable(confusionMatrix(predictedAR, cfxDS$testing$cfx_sr)$byClass[c(1:4, 11)], caption = "Cefoxitin Support Vector Machines Performance Statistics") # SSVMsitivity, Specificity, Pos Pred Value, Neg Pred value, Balanced Accuracy
cfxSVM_tune <- plot.train(SVM$cfxSVM_model) #graphs Tuning Accuracy
#cfxSVM_Imp <- varImp(SVM$cfxSVM_model) %>% dotPlot() #graphs Importance
```


### Ciprofloxacin

#### Training

```{r, echo = FALSE}
confusionMatrix.train(SVM$cipSVM_model)
knitr::kable(SVM$cipSVM_model$bestTune, caption = "CIP Support Vector Machines tuned parameters")
```

#### Final Model AssessmSVMt

```{r, echo=FALSE}
predictedAR <- predict(SVM$cipSVM_model, cipDS$testing) #full predictions
knitr::kable(confusionMatrix(predictedAR, cipDS$testing$cip_sr)$table, caption = "Ciprofloxacin Support Vector Machines Confusion Matrix") #confusion matrix
knitr::kable(confusionMatrix(predictedAR, cipDS$testing$cip_sr)$byClass[c(1:4, 11)], caption = "Ciprofloxacin Support Vector Machines Performance Statistics") # SSVMsitivity, Specificity, Pos Pred Value, Neg Pred value, Balanced Accuracy
cipSVM_tune <- plot.train(SVM$cipSVM_model) #graphs Tuning Accuracy
#cipSVM_Imp <- varImp(SVM$cipSVM_model) %>% dotPlot() #graphs Importance
```

### Hyperparameter Tuning Comparison

```{r, echo = FALSE, fig.cap="ROC vs Cost: AZM, CFX, and CIP from top to bottom"}
gridExtra::grid.arrange(azmSVM_tune, cfxSVM_tune, cipSVM_tune)
```

## Extreme Gradient Boosting

### Azithromycin

#### Training

```{r, echo = FALSE}
confusionMatrix.train(XGB$azmXGB_model)
knitr::kable(XGB$azmXGB_model$bestTune, caption = "AZM Extreme Gradient Boosting tuned parameters")
```

#### Final Model AssessmXGBt

```{r, echo=FALSE, warning=FALSE}
predictedAR <- predict(XGB$azmXGB_model, azmDS$testing) #full predictions
knitr::kable(confusionMatrix(predictedAR, azmDS$testing$azm_sr)$table, caption = "Azithromycin Extreme Gradient Boost Confusion Matrix") #confusion matrix
knitr::kable(confusionMatrix(predictedAR, azmDS$testing$azm_sr)$byClass[c(1:4, 11)], caption = "Azithromycin Extreme Gradient Boost Performance Statistics") # SXGBsitivity, Specificity, Pos Pred Value, Neg Pred value, Balanced Accuracy
azmXGB_tune <- plot.train(XGB$azmXGB_model) #graphs Tuning Accuracy
azmXGB_Imp <- varImp(XGB$azmXGB_model) %>% dotPlot() #graphs Importance
```

### Cefoxitin

#### Training

```{r, echo = FALSE}
confusionMatrix.train(XGB$cfxXGB_model)
knitr::kable(XGB$cfxXGB_model$bestTune, caption = "CFX Extreme Gradient Boosting tuned parameters")
```

#### Final Model AssessmXGBt

```{r, echo=FALSE, warning=FALSE}
predictedAR <- predict(XGB$cfxXGB_model, cfxDS$testing) #full predictions
knitr::kable(confusionMatrix(predictedAR, cfxDS$testing$cfx_sr)$table, caption = "Cefoxitin Extreme Gradient Boost Confusion Matrix") #confusion matrix
knitr::kable(confusionMatrix(predictedAR, cfxDS$testing$cfx_sr)$byClass[c(1:4, 11)], caption = "Cefoxitin Extreme Gradient Boost Performance Statistics") # SXGBsitivity, Specificity, Pos Pred Value, Neg Pred value, Balanced Accuracy
cfxXGB_tune <- plot.train(XGB$cfxXGB_model) #graphs Tuning Accuracy
cfxXGB_Imp <- varImp(XGB$cfxXGB_model) %>% dotPlot() #graphs Importance
```


### Ciprofloxacin

#### Training

```{r, echo = FALSE}
confusionMatrix.train(XGB$cipXGB_model)
knitr::kable(XGB$cipXGB_model$bestTune, caption = "CIP Extreme Gradient Boosting tuned parameters")
```

#### Final Model AssessmXGBt

```{r, echo=FALSE, warning=FALSE}
predictedAR <- predict(XGB$cipXGB_model, cipDS$testing) #full predictions
knitr::kable(confusionMatrix(predictedAR, cipDS$testing$cip_sr)$table, caption = "Ciprofloxacin Extreme Gradient Boost Confusion Matrix") #confusion matrix
knitr::kable(confusionMatrix(predictedAR, cipDS$testing$cip_sr)$byClass[c(1:4, 11)], caption = "Ciprofloxacin Extreme Gradient Boost Performance Statistics") # SXGBsitivity, Specificity, Pos Pred Value, Neg Pred value, Balanced Accuracy
cipXGB_tune <- plot.train(XGB$cipXGB_model) #graphs Tuning Accuracy
cipXGB_Imp <- varImp(XGB$cipXGB_model) %>% dotPlot() #graphs Importance
```

### Hyperparameter Tuning Comparison

```{r, echo = FALSE, fig.cap = "ROC vs Boosting Iterations: AZM, CFX, and CIP from top to bottom"}
azmXGB_tune
cfxXGB_tune
cipXGB_tune
```

## Random Forest

### Azithromycin

#### Training

```{r, echo = FALSE}
confusionMatrix.train(RF$azmRF_model)
knitr::kable(RF$azmRF_model$bestTune, caption = "AZM Random Forest tuned parameters")
```

#### Final Model AssessmRFt

```{r, echo=FALSE}
predictedAR <- predict(RF$azmRF_model, azmDS$testing) #full predictions
knitr::kable(confusionMatrix(predictedAR, azmDS$testing$azm_sr)$table, caption = "Azithromycin Random Forest Confusion Matrix") #confusion matrix
knitr::kable(confusionMatrix(predictedAR, azmDS$testing$azm_sr)$byClass[c(1:4, 11)], caption = "Azithromycin Random Forest Performance Statistics") # SRFsitivity, Specificity, Pos Pred Value, Neg Pred value, Balanced Accuracy
azmRF_tune <- plot.train(RF$azmRF_model) #graphs Tuning Accuracy
azmRF_Imp <- varImp(RF$azmRF_model) %>% dotPlot() #graphs Importance
```

### Cefoxitin

#### Training

```{r, echo = FALSE}
confusionMatrix.train(RF$cfxRF_model)
knitr::kable(RF$cfxRF_model$bestTune, caption = "CFX Random Forest tuned parameters")
```

#### Final Model AssessmRFt

```{r, echo=FALSE}
predictedAR <- predict(RF$cfxRF_model, cfxDS$testing) #full predictions
knitr::kable(confusionMatrix(predictedAR, cfxDS$testing$cfx_sr)$table, caption = "Cefoxitin Random Forest Confusion Matrix") #confusion matrix
knitr::kable(confusionMatrix(predictedAR, cfxDS$testing$cfx_sr)$byClass[c(1:4, 11)], caption = "Cefoxitin Random Forest Performance Statistics") # SRFsitivity, Specificity, Pos Pred Value, Neg Pred value, Balanced Accuracy
cfxRF_tune <- plot.train(RF$cfxRF_model) #graphs Tuning Accuracy
cfxRF_Imp <- varImp(RF$cfxRF_model) %>% dotPlot() #graphs Importance
```


### Ciprofloxacin

#### Training

```{r, echo = FALSE}
confusionMatrix.train(RF$cipRF_model)
knitr::kable(RF$cipRF_model$bestTune, caption = "CIP Random Forest tuned parameters")
```

#### Final Model AssessmRFt

```{r, echo=FALSE}
predictedAR <- predict(RF$cipRF_model, cipDS$testing) #full predictions
knitr::kable(confusionMatrix(predictedAR, cipDS$testing$cip_sr)$table, caption = "Ciprofloxacin Random Forest Confusion Matrix") #confusion matrix
knitr::kable(confusionMatrix(predictedAR, cipDS$testing$cip_sr)$byClass[c(1:4, 11)], caption = "Ciprofloxacin Random Forest Performance Statistics") # SRFsitivity, Specificity, Pos Pred Value, Neg Pred value, Balanced Accuracy
cipRF_tune <- plot.train(RF$cipRF_model) #graphs Tuning Accuracy
cipRF_Imp <- varImp(RF$cipRF_model) %>% dotPlot() #graphs Importance
```

### Hyperparameter Tuning Comparison

```{r, echo = FALSE, fig.cap = "ROC vs Amount of Predictors: AZM, CFX, and CIP from top to bottom"}
gridExtra::grid.arrange(azmRF_tune, cfxRF_tune, cipRF_tune)
```

## Parameter Importance

### Azithromycin

```{r, echo = FALSE, fig.cap = "AZM Predictor Importance Comparison"}
ggarrange(azmEN_Imp, azmXGB_Imp, azmRF_Imp,labels = c("EN", "XGB", "RF"))
```


### Cefoxitin

```{r, echo = FALSE, fig.cap = "CFX Predictor Importance Comparison"}
ggarrange(cfxEN_Imp, cfxXGB_Imp, cfxRF_Imp,labels = c("EN", "XGB", "RF"))
```

### Ciprofloxacin

```{r, echo = FALSE, fig.cap = "CIP Predictor Importance Comparison"}
ggarrange(cipEN_Imp, cipXGB_Imp, cipRF_Imp,labels = c("EN", "XGB", "RF"))
```

# Discussion