---
title: "R Notebook"
output: html_notebook
---

#Reference code tutorial:
#uc-r.github.io/svm



```{r}
library(tidyverse)
library(readr)

# read metadata.csv with only "Sample_ID" and "*_sr" features
metadata <- read_csv("./metadata.csv") %>% 
  select(Sample_ID,(matches("_sr$")))

# read antibiotic resistance associated unitigs
unitigs_azm <- read_table2("azm_sr_gwas_filtered_unitigs.Rtab")
unitigs_cfx <- read_table2("cfx_sr_gwas_filtered_unitigs.Rtab")
unitigs_cip <- read_table2("cip_sr_gwas_filtered_unitigs.Rtab")

#Function to transpose unitigs table, join metadata, then create .csv file
unitigs <- function(unitig){
  
  # deparse + substitute combo to extract object name
  unitig_name <- deparse(substitute(unitig))
  
  '%>%' <- tidyr::'%>%'
  
  unitig %>%
    # transpose
    tidyr::pivot_longer(-pattern_id, "Sample_ID", "value") %>% 
    tidyr::pivot_wider(Sample_ID, pattern_id) %>%
    # join relevant antibiotic metadata
    dplyr::inner_join(dplyr::select(metadata, "Sample_ID", dplyr::matches(stringr::str_extract(unitig_name, "[[:alpha:]]+$"))), by = "Sample_ID") %>% 
    dplyr::select(dplyr::last_col(), dplyr::everything(), -"Sample_ID") %>% 
    # rename unitigs to format "geneX" where X is a positive integer
    data.table::setnames(., old = names(.[2:length(.)]), new = paste0('gene', seq_along(.[2:length(.)])), skip_absent = T) %>% 
    tidyr::drop_na() %>% 
    # create .csv file
    readr::write_csv(path=paste0('./', stringr::str_extract(unitig_name,"[[:alpha:]]+[_][[:alpha:]]+"), '.csv'))
}

# Manual Transpose, join, and create .csv files of azm unitigs
# unitigs_azm %>%
#   pivot_longer(-pattern_id, "Sample_ID", "value") %>%
#   pivot_wider(Sample_ID, pattern_id) %>%
#   inner_join(select(metadata, "Sample_ID", matches('^azm')), by = "Sample_ID") %>%
#   select('Sample_ID', last_col(), everything()) %>%
#   data.table::setnames(., old = names(.[2:length(.)]), new = paste0('gene', seq_along(.[2:length(.)])), skip_absent = T) %>% 
#   drop_na() %>%
#   write_csv(path="./unitigs_azm.csv")

# Use 'unitigs' function to create .csv files
unitigs(unitigs_azm)
unitigs(unitigs_cfx)
unitigs(unitigs_cip)
```



#Split into training and testing:
```{r}
createDS <- function(csvname, percent=0.7, seed=05101929){
  stopifnot(is.character(csvname))
  
  '%>%' <- tidyr::'%>%'
  
#Import dataset
svmData <- readr::read_csv(csvname, col_types = readr::cols(.default = readr::col_factor(NULL))) %>% 
  dplyr::select_if(sapply(., nlevels) > 1)
#Split
set.seed(seed)
index <- dplyr::select(svmData,dplyr::matches(stringr::str_extract(unlist(stringr::str_split(csvname, "_"))[2],"^[[:alpha:]]+"))) %>% 
  unlist() %>% 
  caret::createDataPartition(times = 1, p = percent, list = FALSE)

dfTrain <- svmData %>% dplyr::slice(index) #training dataset
dfTest <- svmData %>% dplyr::slice(-index) #testing dataset

#return datasets
dataSets <- list(training = dfTrain, testing = dfTest)
invisible(dataSets)
}
```

#Create dataframes, view:

```{r}
azmDS <- createDS("./unitigs_azm.csv")
tibble::glimpse(azmDS$training[1:6])
```



#Load in SVM-specific packages:

```{r, results='hide'}
library(kernlab)
library(e1071)
library(ISLR)
library(RColorBrewer)
```

#Test SVM model 1: linear assumption

```{r}
svmfit1 <- svm(azm_sr~., data=azmDS$training, kernel="linear", scale=FALSE)
svmfit1
```


#Tune the 'cost' parameter to identify optimal misclassification threshold:
```{r}
tune.out1 <- tune(svm, azm_sr~., data=azmDS$training, kernel = "linear", ranges=list(cost = c(0.001, 0.01, 0.1, 1, 5, 10, 100)))
(bestmod1 <- tune.out1$best.model)
```


#Low cost identified

#Construct confusion matrix:

```{r}
ypred1 <- predict(bestmod1, azmDS$training)
(misclass1 <- table(predict = ypred1, truth = azmDS[[1]][[1]]))
```


#In this model, 
```{r}
(290+2085)/(290+2085+23+37)*100
```

#... percent of observations were identified correctly.



#Moving on to the testing set:

```{r}
svmfit2 <- svm(azm_sr~., data=azmDS$testing, kernel="linear", scale=FALSE)
svmfit2
```

#Verify cost function:
```{r}
tune.out2 <- tune(svm, azm_sr~., data=azmDS$testing, kernel = "linear", ranges=list(cost = c(0.001, 0.01, 0.1, 1, 5, 10, 100)))
(bestmod2 <- tune.out2$best.model)
```

#Same cost parameter.

```{r}
ypred2 <- predict(bestmod2, azmDS$testing)
(misclass2 <- table(predict = ypred2, truth = azmDS[[2]][[1]]))
```


#And in the testing set,
```{r}
(127+895)/(127+895+14+7)*100
```

#... percent of observations were identified correctly.


####.
####Next model set:
####CIP
####.



```{r}
cipDS <- createDS("./unitigs_cip.csv")
tibble::glimpse(cipDS$training[1:6])
```


#Model 1: Linear assumption

```{r}
cipfit1 <- svm(cip_sr~., data=cipDS$training, kernel="linear", scale=FALSE)
cipfit1
```


#Determing optimal misclassification cost:
```{r}
tune.out3 <- tune(svm, cip_sr~., data=cipDS$training, kernel = "linear", ranges=list(cost = c(0.001, 0.01, 0.1, 1, 5, 10, 100)))
(bestmod3 <- tune.out3$best.model)
```



